The Design and Implementation of the Synchronous Language \CEU

### 2016-10-07 paper.tex
(>) Comments Noemi/Adriano

\CEU aims to offer a simpler semantics than Esterel with more determinism and 
fine-grained control over program execution.
%
The following list summarizes the contributions of our design in this 
direction:
%
\begin{itemize}
%
\item \emph{Event-triggered notion of time bound to the semantics of the 
language.}
Event-driven programming is popular in many domains, such as server and GUI 
development.
We believe that programmers are more familiar with dealing with events in 
isolation, simplifying the reasoning about concurrency.
In addition, the uniqueness of external events is a prerequisite for the 
concurrency checks of \CEU.
%
\item \emph{Deterministic intra-reaction execution and communication.}
Determinism in \CEU does not depend on additional levels of static analysis.
It encompasses the whole language, including memory accesses, system calls, and 
stack-based internal events.
Programmers can figure out which statement executes next following clear
sequential rules.
%
\item \emph{Static concurrency checks.}
Although execution is deterministic, the \CEU compiler still advises about 
suspicious statements that can react concurrently to the same event. 
%appropriate serialization.
%
\item \emph{Safe integration with \C.}
When dealing with concurrent system calls, programmers can provide annotations 
to reduce false positives in the static checks, or to force non commutative 
concurrent behavior.
\CEU also requires finalization clauses to handle pointers representing 
resources.
%
\item \emph{First-class timers with dedicated syntax and automatic 
synchronization.}
Given the omnipresence of timers in embedded systems, a dedicated syntax can 
simplify the development and readability of programs.
Furthermore, automatic synchronization releases the programmer from the burden 
of adjusting timers in sequence and in parallel.
%
\end{itemize}

Our synchronous and deterministic approach also leads to some limitations as 
follows:
%
\begin{itemize}
\item \emph{Execution is purely reactive as result of event-triggered 
reactions.}
Since only event occurrences can start reactions, programs cannot execute 
proactively in the absence of events.
In addition, \code{await} statements cannot awake in the same reaction they are
reached.
\item \emph{Reactions must execute in bounded time due to the synchronous 
hypothesis.}
As a synchronous language, \CEU requires CPU times for reactions to be 
negligible in comparison to the rate of incoming events.
\item \emph{Execution is sequential because of intra-reaction determinism.}
%We discuss some implementation considerations in 
%Section~\ref{sec.impl.single}.
The deterministic semantics of \CEU does not make implicit parallelization easy 
(to be discussed in Section~\ref{sec.impl.single}).
\end{itemize}
%
Nonetheless, we advocate keeping a tractable synchronous reactive core with 
support for shared memory concurrency and deterministic execution.
%
To deal with the limitations above, we recommend memory-isolated parallelizable 
asynchronous primitives as separate extensions to the synchronous 
core~\cite{esterel.crp} (which are not in the scope of this paper).

>>>>

\CEU aims to offer a simple semantics with internal determinism and
fine-grained control over execution.
%
The following list summarizes the contributions of our design in this 
direction:
%
\begin{itemize}
%
\item \emph{Event-triggered notion of time bound to the semantics of the 
language.}
Event-driven programming is popular in many domains, such as server and GUI 
development.
We believe that programmers are more familiar with dealing with events in 
isolation, simplifying the reasoning about concurrency.
In addition, the uniqueness of external events is a prerequisite for the 
concurrency checks of \CEU.
%
\item \emph{Deterministic intra-reaction execution and communication.}
Determinism in \CEU does not depend on additional levels of static analysis.
It encompasses the whole language, including memory accesses, system calls, and 
stack-based internal events.
Programmers can figure out which statement executes next following clear
sequential rules.
%
\item \emph{Static concurrency checks.}
Although execution is deterministic, the \CEU compiler still advises about 
suspicious statements that can react concurrently to the same event. 
%appropriate serialization.
%
\item \emph{Safe integration with \C.}
When dealing with concurrent system calls, programmers can provide annotations 
to reduce false positives in the static checks, or to force non commutative 
concurrent behavior.
\CEU also requires finalization clauses to handle pointers representing 
resources.
%
\item \emph{First-class timers with dedicated syntax and automatic 
synchronization.}
Given the demand of timers in embedded systems, a dedicated syntax can simplify
the development and readability of programs.
Furthermore, automatic synchronization releases the programmer from the burden 
of adjusting timers in sequence and in parallel.
%
\end{itemize}

Our approach for synchronous and deterministic execution also leads to some
limitations as follows:
%
\begin{itemize}
\item \emph{Execution is purely reactive as result of event-triggered 
reactions.}
Since only event occurrences can start reactions, programs cannot execute 
proactively in the absence of events.
In addition, \code{await} statements cannot awake in the same reaction they are
reached.
\item \emph{Reactions must execute in bounded time due to the synchronous 
hypothesis.}
As a synchronous language, \CEU requires CPU times for reactions to be 
negligible in comparison to the rate of incoming events.
\item \emph{Execution is sequential because of intra-reaction determinism.}
%We discuss some implementation considerations in 
%Section~\ref{sec.impl.single}.
The deterministic semantics of \CEU does not make implicit parallelization easy 
(to be discussed in Section~\ref{sec.impl.single}).
\end{itemize}
%
Nonetheless, we advocate keeping a tractable synchronous reactive core with 
support for shared memory concurrency and deterministic execution.
%
To deal with the limitations above, we recommend memory-isolated parallelizable 
asynchronous primitives as separate extensions to the synchronous 
core~\cite{esterel.crp} (which are not in the scope of this paper).
